### 2024-09-20 other/literature/linear_model.qmd

---
title: "Linear model"
author: "Max LI"
date: "2024/9/20"
format: pdf
---

## Model Summary

### Formula

The model predicts the number of crimes (`num_crimes`) based on the number of mental health apprehensions (`num_mha_reports`) using a linear relationship.

### Equation

$$
C = 6748.244 + 2.451 \times MHA
$$

### Coefficients and Interpretation

-   **C**\
    C stands for the predicted number of crimes, which is also the dependent variable in this model.

-   **Intercept:** 6748.244\
    This suggests that, even when there are no MHA reports, the baseline number of crimes in a division is approximately 6748.

-   **MHA:** 2.451\
    For each additional MHA report, the model predicts an increase of approximately 2.45 crimes in the corresponding division.

### Interpretation

The model suggests a positive linear relationship between MHA reports and crimes, meaning that as MHA reports increase, the number of crimes tends to increase as well.